---
title: "Getting Started with unifyr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with unifyr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`unifyr` provides tools for auditing, validating, and preparing panel datasets before statistical analysis.

Panel datasets often contain duplicate unit-time observations, missing time periods, irregular gaps, and imbalance. These issues can affect fixed effects models, difference-in-differences designs, event studies, and other panel-data methods.

The goal of `unifyr` is to help users identify these issues before estimation.

# Load the package

```{r}
library(unifyr)
```

# Example panel dataset

`unifyr` includes a small example dataset called `example_panel`.

```{r}
data(example_panel)

example_panel
```

The dataset intentionally includes:

- a duplicate unit-time observation
- missing unit-time cells
- an unbalanced panel structure

This makes it useful for demonstrating panel-data diagnostics.

# Audit the panel

The main function is `audit_panel()`.

```{r}
audit_panel(example_panel, id = id, time = year)
```

This gives a quick overview of the panel structure, including whether the panel is balanced and whether there are missing or duplicate unit-time cells.

# Find duplicate observations

Duplicate unit-time observations are a common problem in panel datasets.

```{r}
duplicate_summary(example_panel, id = id, time = year)
```

# Summarize gaps

`gap_summary()` identifies missing time periods by panel unit.

```{r}
gap_summary(example_panel, id = id, time = year)
```

# Flag row-level issues

`flag_panel_issues()` adds diagnostic flags to the data.

```{r}
flag_panel_issues(example_panel, id = id, time = year)
```

# Complete a panel grid

`complete_panel()` creates a complete unit-time grid. It does not impute missing outcome values.

Because `complete_panel()` requires unique unit-time cells, we first remove duplicate id-time observations from the example dataset.

```{r}
example_panel_unique <- example_panel |>
  dplyr::distinct(id, year, .keep_all = TRUE)

complete_panel(example_panel_unique, id = id, time = year)
```

# Typical workflow

A typical `unifyr` workflow is:

```{r, eval = FALSE}
library(unifyr)

audit_panel(my_data, id = unit_id, time = year)

duplicate_summary(my_data, id = unit_id, time = year)

gap_summary(my_data, id = unit_id, time = year)

clean_data <- my_data |>
  dplyr::distinct(unit_id, year, .keep_all = TRUE)

complete_panel(clean_data, id = unit_id, time = year)
```

# Summary

`unifyr` is designed to provide a transparent and reproducible workflow for panel-data quality assurance.

Use it before fitting panel models, difference-in-differences designs, event studies, or other longitudinal-data analyses.
